The Private School Network: Recruiting Visits to Private High Schools by Public and Private Universities

### 2023-07-25 text/manuscript_rihe.Rmd
added rihe manuscript

---
title: "The Private School Network: Recruiting Visits to Private High Schools by Public and Private Universities"
author: 
  - Ozan Jaquette
  - Crystal Han
  - Irma Castañeda
bibliography: ./../other/burd_recruiting_chapter.bib
output: 
  bookdown::pdf_document2:
    toc: FALSE
csl: ./../other/apa.csl
urlcolor: blue
header-includes:
      - \usepackage{pdflscape}
      - \usepackage{geometry}
      - \usepackage{floatrow}
      - \usepackage{setspace}
      #- \doublespacing
      - \floatsetup{capposition=top}
fontsize: 11pt
---

@@

@@

Prior to the pandemic, admissions counselors from colleges and universities traveled to local high schools each year in order to meet prospective students and maintain relationships with guidance counselors at feeder schools [@RN3519]. @RN4721 analyzed off-campus recruiting visits made in 2017 by 15 public research universities. Surprisingly, 12 of the 15 universities made more visits to *out-of-state* high schools than *in-state* high schools. These out-of-state visits focused on public schools in affluent, predominantly white communities. Additionally, most public research universities in our sample made a disproportionate number of visits to out-of-state private high schools. For the University of California-Berkeley, for example, 155 (43%) of the 360 out-of-state high school visits were to private high schools. For the University of Colorado-Boulder, 321 (36%) of the 901 out-of-state high school visits were to private high schools.

@@

Whereas @RN4721 primarily analyzed visits to public high schools, this chapter focuses on visits to private high schools. More specifically, we investigate how visits to private high schools by public research universities compare to visits to private high schools by selective private universities.

@@

The mission of selective private colleges and universities is more squarely associated with social reproduction rather than social mobility. Historically, these institutions depend on patronage from philanthropic interests -- in connection with business and religious interests -- and on wealthy households who provide tuition revenue and donations [@RN4517]. In turn, selective private institutions contribute to social reproduction by providing access to their patrons. The "Chetty" data on parental income by college show that the overwhelming majority of students at selective private institutions have parental income in the top quintile [@RN4722; @RN4723]. A significant caveat is that religiously affiliated universities may prioritize enrollment opportunities for students of the same faith regardless of wealth, but even here the mission is to benefit families from a particular cultural status group. <!-- [CITE] -->

@@

Analyses of recruiting visits to private high schools can contribute novel insights about the "privatization" of public research universities. Within the scholarly literature on privatization, one strand of scholarship conceptualizes privatization as a process whereby declines in government funding cause public institutions to grow alternative revenue sources, particularly revenue from tuition and research commercialization [@RN4543; @RN2101]. Of relevance to this chapter, public research universities responded to declines in state appropriations by growing enrollment from nonresident students [@RN3753], who typically pay two- to three-times more than resident students. A second strand within the privatization literature conceptualizes privatization as a process whereby public institutions become more similar -- in terms of characteristics or behavior -- to for-profit or private non-profit organizations [@RN4543]. A weakness of this literature is that few empirical studies analyze public institutions on a behavior associated with private institutions.

Therefore, this chapter compares public research universities to selective private universities on the dimension of recruiting visits to private high schools, a behavior associated with the social reproduction orientation of selective private universities. Our primary goal is to learn about public research universities, using the behavior of private universities as a benchmark to develop more thoughtful insights about public research universities.  We analyze recruiting visits made in the 2017 calendar year by a sample of 15 public research universities and 14 selective private universities. <!-- and 12 selective private liberal arts colleges. -->

@@

Our broader project collected data on off-campus recruiting visits made by a convenience sample of colleges and universities during the 2017 calendar year. The data collection sample comes from two different lists of postsecondary institutions: all public research-extensive universities as defined by the 2000 Carnegie Classification (N=102), and all private universities in the top 100 of U.S. News and World Report National Universities rankings (N=58). <!-- and all private colleges in the top 50 of U.S. News and World Report Liberal Arts Colleges rankings (N=47).-->For each of these institutions, we investigated their admissions website for pages that posted their upcoming off-campus recruiting visits. For institutions that posted such pages, we scraped the pages once per week throughout 2017. Many colleges and universities only posted certain kinds of events (e.g., hotel receptions and national college fairs) but not others (e.g., day-time visits to high schools). These institutions are excluded from the analyses. Our final analysis sample consists of 15 public research universities and 14 private research universities. <!--12 private liberal arts colleges [MAYBE ADD TABLE/FIGURE OR APPENDIX TABLE/FIGURE ON CHARACTERISTICS OF INSTITUTIONS IN OUR SAMPLE?]-->

@@

@@

Appendix Figures \@ref(fig:events-count-pubu) and \@ref(fig:events-count-privu) show the number of visits by "type" (public high school, private high school, community college, other) for public research universities and selective private universities, respectively. On average, the `r univ_sample_df %>% filter(classification == 'public_research') %>% count() %>% as.numeric()` public research universities in our sample conducted `r round((events_df %>% filter(classification == 'public_research') %>% count() %>% as.numeric())/(univ_sample_df %>% filter(classification == 'public_research') %>% count() %>% as.numeric()), digits = 0)` off-campus recruiting events in 2017, ranging from a low of `r events_df %>% filter(univ_id == '139959') %>% count() %>% as.numeric()` for the University of Georgia to a high of `r events_df %>% filter(univ_id == '100751') %>% count() %>% as.numeric()` for the University of Alabama. By contrast, the `r univ_sample_df %>% filter(classification == 'private_national') %>% count() %>% as.numeric()` selective private universities in our sample conducted an average of `r round((events_df %>% filter(classification == 'private_national') %>% count() %>% as.numeric())/(univ_sample_df %>% filter(classification == 'private_national') %>% count() %>% as.numeric()), digits = 0)` events, ranging from a low of `r events_df %>% filter(univ_id == '186867') %>% count() %>% as.numeric()` by Stevens Institute of Technology (located in Hoboken, NJ) to a high of `r events_df %>% filter(univ_id == '239105') %>% count() %>% as.numeric()` by Marquette University (located in Milwaukee, WI).  Across all public research universities, `r round((events_df %>% filter(classification == 'public_research', event_type %in% c('pub_hs','priv_hs')) %>% count() %>% as.numeric())/(events_df %>% filter(classification == 'public_research') %>% count() %>% as.numeric())*100, digits=1)`% of visits were to a public or private high school, whereas for selective private universities `r round((events_df %>% filter(classification == 'private_national', event_type %in% c('pub_hs','priv_hs')) %>% count() %>% as.numeric())/(events_df %>% filter(classification == 'private_national') %>% count() %>% as.numeric())*100, digits=1)`% of all visits were to a public or private high school. The lower percentage for public research universities is because many public research universities made a substantial number of visits to community colleges whereas private universities did not.

Figures \@ref(fig:events-hs-count-pubu) and \@ref(fig:events-hs-count-privu) show the number of visits to private high schools and public high schools for public research universities (Figure \@ref(fig:events-hs-count-pubu)) and selective private universities (Figure \@ref(fig:events-hs-count-privu)), with each bar shaded by percent in-state versus out-of-state visits. Only three of the 15 public research universities -- UC-Berkeley, UC-San Diego, and UC-Riverside -- made more in-state than out-of-state total high school visits. Notably, the University of California voluntarily adopted a nonresident enrollment cap in 2017, following pressure from state legislators who had issued a bill to cap nonresident enrollment. Of the 14 selective private universities in our sample, only Baylor University (Waco, TX) made more in-state than out-of-state total high school visits.

@@

Appendix X describes how we constructed social network data. Briefly, we began by constructing a two-mode social network of visits to private high schools (mode 1) by universities (mode 2). Each university and each private high school is a "vertex." A visit from a university to a school is an "edge." Because large two-mode networks like our school-university network are difficult to visualize, two-mode networks are often analyzed as one-mode networks [@RN2457;@RN4663]. We transformed our two-mode school-university network into a one-mode university network in which each university is a vertex and two universities share an edge if both universities visited at least one high school in common. The *weight* of an edge indicates the number of high schools that both universities visited. For example, if there are 200 private high schools that received a visit from both Notre Dame and Villanova, the weight of the edge between Notre Dame and Villanova is 200. We created three separate one-mode university networks: a public university network; a private university network; a public and private university network.

@@

@@

# Social network data appendix

>>>>

---
title: "The Private School Network: Recruiting Visits to Private High Schools by Public and Private Universities"
#author: 
#  - Ozan Jaquette
#  - Crystal Han
#  - Irma Castañeda
bibliography: ./../other/burd_recruiting_chapter.bib
output: 
  bookdown::pdf_document2:
    toc: FALSE
    pandoc_args: !expr rmdfiltr::add_wordcount_filter(rmdfiltr::add_citeproc_filter(args = NULL))
csl: ./../other/apa.csl
urlcolor: blue
header-includes:
      - \usepackage{pdflscape}
      - \usepackage{geometry}
      - \usepackage{floatrow}
      - \usepackage{setspace}
      #- \doublespacing
      - \floatsetup{capposition=top}
fontsize: 12pt
---

@@

\setstretch{1.5}

ABSTRACT

Few empirical analyses of privatization compare public and private universities on a behavior associated with private universities. This manuscript analyzes off-campus recruiting visits to private high schools, which were made in 2017 by a convenience sample of 15 public research universities and 14 selective, private universities. Sociological scholarship on the relationship between high schools and universities conceives of off-campus recruiting visits as an indicator of a social relation. Therefore, we utilize social network concepts and methods to compare the recruiting networks of public and private universities. With respect to scale (research question 1), most public research universities in our sample made a disproportionate number of visits to private (out-of-state) high schools, though not as many as private universities. With respect to characteristics of visited schools (RQ2), both public and private universities tended to visit predominantly white schools in their home geographic region and the South. Public universities tended to visit lower ranked private high schools than private universities. Surprisingly, several public research universities visited sectarian private high schools at a rate similar to sectarian private universities. With respect to overlap (RQ3), descriptive statistics and community detection methods revealed substantial overlap in the recruiting networks of public and private universities.

@@

In the early twenty-first century, "privatization" received substantial attention from scholars of higher education [e.g., @RN921; @RN2101; @RN1118; @RN1577; @RN2111]. @RN4543 argue that scholarship on privatization waned because of a lack of conceptual and empirical clarity. One strand of scholarship conceptualizes privatization as a process whereby declines in government funding cause public institutions to grow alternative revenue sources. Public research universities responded, for example, by pursuing research commercialization and nonresident tuition revenue, and donative resources [e.g., @RN3753; @RN2101; @RN1644]. A second strand within the privatization literature conceptualizes privatization as a process whereby public institutions become more similar -- in terms of characteristics or behavior -- to for-profit or private non-profit organizations [@RN4543]. A weakness of this literature is that few empirical studies analyze public institutions on a behavior associated with private institutions. Moreover, few studies directly compare public and private institutions on the same behavioral dimension.

We argue that scholarship on recruiting behavior can shed light on the extent to which public universities are engaging in behaviors associated with private universities. Admissions counselors from colleges and universities visit local high schools in order to meet prospective students and maintain relationships with guidance counselors at feeder schools [@RN3519]. @RN4758 analyzed off-campus recruiting visits made in 2017 by 15 public research universities, finding that 12 of these universities made more visits to *out-of-state* high schools than *in-state* high schools. These out-of-state visits focused on public schools in affluent, predominantly white communities. Whereas @RN4758 primarily analyzed visits to public high schools, this chapter focuses on visits to private high schools. More specifically, we investigate how visits to private high schools by public research universities compare to visits to private high schools by selective private universities.

Our focus on recruiting visits to private high schools is motivated by seminal scholarship within the sociology of education. Sociologist Max Weber is credited for the insight that education has a "dual character," on one hand, serving *social mobility* by redistributing access to privileged positions and, on the other, serving *social reproduction* by granting privileged families disproportionate access to the educational institutions that confer pathways to privileged positions [@RN2377; @RN1133].

The mission of selective private colleges and universities is more squarely associated with social reproduction rather than social mobility. Historically, these institutions depend on patronage from philanthropic interests -- in connection with business and religious interests -- and on wealthy households who provide tuition revenue and donations [@RN4517]. In turn, selective private institutions contribute to social reproduction by providing access to their patrons. The "Chetty" data on parental income by college show that the overwhelming majority of students at selective private institutions have parental income in the top quintile [@RN4722; @RN4723]. A significant caveat is that religiously affiliated universities may prioritize enrollment opportunities for students of the same faith regardless of wealth, but even here the mission is to benefit families from a particular cultural status group [@RN4517].

This manuscript compares public research universities to selective private universities on the dimension of recruiting visits to private high schools, a behavior associated with the social reproduction orientation of selective private universities. Our primary goal is to learn about public research universities, using the behavior of private universities as a benchmark to develop more thoughtful insights about public research universities. We analyze recruiting visits made in the 2017 calendar year by a sample of 15 public research universities and 14 selective private universities. 
<!-- and 12 selective private liberal arts colleges. -->

@@

Little is known about recruiting at private high schools by public universities. Therefore, we conduct descriptive analyses that seek to identify basic patterns and relationships. Because a recruiting visit to the same private high school is an indicator of a relationship between two actors, we also utilize social network methods in order to compare the recruiting networks of public universities to those of private universities. Analyses are informed by the following three research questions:

The remainder of the manuscript is organized as follows. First, we situate off-campus recruiting visits within the broader process of recruiting students and we review scholarship on the relationship between high schools and universities, focusing on scholarship from sociology. This literature motivates off-campus recruiting visits as an indicator of a social relation. Second, we introduce concepts from social network analysis and motivate empirical research questions. Third, we describe data and methods. Fourth, we present results. Finally, we discuss implications for scholarship.

<!--  ORIGINAL INTRODUCTION

Prior to the pandemic, admissions counselors from colleges and universities traveled to local high schools each year in order to meet prospective students and maintain relationships with guidance counselors at feeder schools [@RN3519]. @RN4758 analyzed off-campus recruiting visits made in 2017 by 15 public research universities. Surprisingly, 12 of the 15 universities made more visits to *out-of-state* high schools than *in-state* high schools. These out-of-state visits focused on public schools in affluent, predominantly white communities. Additionally, most public research universities in our sample made a disproportionate number of visits to out-of-state private high schools. For the University of California-Berkeley, for example, 155 (43%) of the 360 out-of-state high school visits were to private high schools. For the University of Colorado-Boulder, 321 (36%) of the 901 out-of-state high school visits were to private high schools.

Whereas @RN4758 primarily analyzed visits to public high schools, this chapter focuses on visits to private high schools. More specifically, we investigate how visits to private high schools by public research universities compare to visits to private high schools by selective private universities.

@@

The mission of selective private colleges and universities is more squarely associated with social reproduction rather than social mobility. Historically, these institutions depend on patronage from philanthropic interests -- in connection with business and religious interests -- and on wealthy households who provide tuition revenue and donations [@RN4517]. In turn, selective private institutions contribute to social reproduction by providing access to their patrons. The "Chetty" data on parental income by college show that the overwhelming majority of students at selective private institutions have parental income in the top quintile [@RN4722; @RN4723]. A significant caveat is that religiously affiliated universities may prioritize enrollment opportunities for students of the same faith regardless of wealth, but even here the mission is to benefit families from a particular cultural status group.

@@

Analyses of recruiting visits to private high schools can contribute novel insights about the "privatization" of public research universities. Within the scholarly literature on privatization, one strand of scholarship conceptualizes privatization as a process whereby declines in government funding cause public institutions to grow alternative revenue sources, particularly revenue from tuition and research commercialization [@RN4543; @RN2101]. Of relevance to this chapter, public research universities responded to declines in state appropriations by growing enrollment from nonresident students [@RN3753], who typically pay two- to three-times more than resident students. A second strand within the privatization literature conceptualizes privatization as a process whereby public institutions become more similar -- in terms of characteristics or behavior -- to for-profit or private non-profit organizations [@RN4543]. A weakness of this literature is that few empirical studies analyze public institutions on a behavior associated with private institutions. [FURTHER, FEW STUDIES COMPARE THE BEHAVIOR OF PUBLIC UNIVERSITIES TO THAT OF PRIVATE UNIVERSITIES ON THAT BEHAVIOR.]

Therefore, this chapter compares public research universities to selective private universities on the dimension of recruiting visits to private high schools, a behavior associated with the social reproduction orientation of selective private universities. Our primary goal is to learn about public research universities, using the behavior of private universities as a benchmark to develop more thoughtful insights about public research universities.  We analyze recruiting visits made in the 2017 calendar year by a sample of 15 public research universities and 14 selective private universities.

@@

Our broader project collected data on off-campus recruiting visits made by a convenience sample of colleges and universities during the 2017 calendar year. The data collection sample comes from two different lists of postsecondary institutions: all public research-extensive universities as defined by the 2000 Carnegie Classification (N=102), and all private universities in the top 100 of U.S. News and World Report National Universities rankings (N=58). <!-- and all private colleges in the top 50 of U.S. News and World Report Liberal Arts Colleges rankings (N=47).-->For each of these institutions, we investigated their admissions website for pages that posted their upcoming off-campus recruiting visits. For institutions that posted such pages, we scraped the pages once per week throughout 2017. Many colleges and universities only posted certain kinds of events (e.g., hotel receptions and national college fairs) but not others (e.g., day-time visits fto high schools). These institutions are excluded from the analyses. Our final analysis sample consists of 15 public research universities and 14 private research universities. <!--12 private liberal arts colleges [MAYBE ADD TABLE/FIGURE OR APPENDIX TABLE/FIGURE ON CHARACTERISTICS OF INSTITUTIONS IN OUR SAMPLE?]-->

@@

## Social network methods

Analyses consist of simple descriptive statistics and descriptive social network analyses.

@@

Figures \@ref(fig:events-count-pubu) and \@ref(fig:events-count-privu) show the number of visits by "type" (public high school, private high school, community college, other) for public research universities and selective private universities, respectively. On average, the `r univ_sample_df %>% filter(classification == 'public_research') %>% count() %>% as.numeric()` public research universities in our sample conducted `r round((events_df %>% filter(classification == 'public_research') %>% count() %>% as.numeric())/(univ_sample_df %>% filter(classification == 'public_research') %>% count() %>% as.numeric()), digits = 0)` off-campus recruiting events in 2017, ranging from a low of `r events_df %>% filter(univ_id == '139959') %>% count() %>% as.numeric()` for the University of Georgia to a high of `r events_df %>% filter(univ_id == '100751') %>% count() %>% as.numeric()` for the University of Alabama. By contrast, the `r univ_sample_df %>% filter(classification == 'private_national') %>% count() %>% as.numeric()` selective private universities in our sample conducted an average of `r round((events_df %>% filter(classification == 'private_national') %>% count() %>% as.numeric())/(univ_sample_df %>% filter(classification == 'private_national') %>% count() %>% as.numeric()), digits = 0)` events, ranging from a low of `r events_df %>% filter(univ_id == '186867') %>% count() %>% as.numeric()` by Stevens Institute of Technology (located in Hoboken, NJ) to a high of `r events_df %>% filter(univ_id == '239105') %>% count() %>% as.numeric()` by Marquette University (located in Milwaukee, WI).  Across all public research universities, `r round((events_df %>% filter(classification == 'public_research', event_type %in% c('pub_hs','priv_hs')) %>% count() %>% as.numeric())/(events_df %>% filter(classification == 'public_research') %>% count() %>% as.numeric())*100, digits=1)`% of visits were to a public or private high school, whereas for selective private universities `r round((events_df %>% filter(classification == 'private_national', event_type %in% c('pub_hs','priv_hs')) %>% count() %>% as.numeric())/(events_df %>% filter(classification == 'private_national') %>% count() %>% as.numeric())*100, digits=1)`% of all visits were to a public or private high school. The lower percentage for public research universities is because many public research universities made a substantial number of visits to community colleges whereas private universities did not.

Figures \@ref(fig:events-hs-count-pubu) and \@ref(fig:events-hs-count-privu) address RQ1 directly by showing the number of visits to private high schools and public high schools for public research universities (Figure \@ref(fig:events-hs-count-pubu)) and selective private universities (Figure \@ref(fig:events-hs-count-privu)), with each bar shaded by percent in-state versus out-of-state visits. Only three of the 15 public research universities -- UC-Berkeley, UC-San Diego, and UC-Riverside -- made more in-state than out-of-state total high school visits. Notably, the University of California voluntarily adopted a nonresident enrollment cap in 2017, following pressure from state legislators who had issued a bill to cap nonresident enrollment. Of the 14 selective private universities in our sample, only Baylor University (Waco, TX) made more in-state than out-of-state total high school visits.

@@

As described in the methods section, we began by constructing a two-mode social network of visits to private high schools (mode 1) by universities (mode 2). Each university and each private high school is a "vertex." A visit from a university to a school is an "edge." Because large two-mode networks like our school-university network are difficult to visualize, two-mode networks are often analyzed as one-mode networks [@RN2457;@RN4663]. We transformed our two-mode school-university network into a one-mode university network in which each university is a vertex and two universities share an edge if both universities visited at least one high school in common. The *weight* of an edge indicates the number of high schools that both universities visited. For example, if there are 200 private high schools that received a visit from both Notre Dame and Villanova, the weight of the edge between Notre Dame and Villanova is 200. We created three separate one-mode university networks: a public university network; a private university network; a public and private university network.

@@

# Conflict of Interest Statement

All authors declare that they have no conflicts of interest.

\singlespacing

@@

### 2023-07-25 text/working/manuscript_rihe_v1.Rmd
added rihe manuscript

---
title: "The Private School Network: Recruiting Visits to Private High Schools by Public and Private Universities"
author: 
  - Ozan Jaquette
  - Crystal Han
  - Irma Castañeda
bibliography: ./../other/burd_recruiting_chapter.bib
output: 
  bookdown::pdf_document2:
    toc: FALSE
csl: ./../other/apa.csl
urlcolor: blue
header-includes:
      - \usepackage{pdflscape}
      - \usepackage{geometry}
      - \usepackage{floatrow}
      - \usepackage{setspace}
      #- \doublespacing
      - \floatsetup{capposition=top}
fontsize: 11pt
---

@@

@@

Prior to the pandemic, admissions counselors from colleges and universities traveled to local high schools each year in order to meet prospective students and maintain relationships with guidance counselors at feeder schools [@RN3519]. @RN4721 analyzed off-campus recruiting visits made in 2017 by 15 public research universities. Surprisingly, 12 of the 15 universities made more visits to *out-of-state* high schools than *in-state* high schools. These out-of-state visits focused on public schools in affluent, predominantly white communities. Additionally, most public research universities in our sample made a disproportionate number of visits to out-of-state private high schools. For the University of California-Berkeley, for example, 155 (43%) of the 360 out-of-state high school visits were to private high schools. For the University of Colorado-Boulder, 321 (36%) of the 901 out-of-state high school visits were to private high schools.

@@

Whereas @RN4721 primarily analyzed visits to public high schools, this chapter focuses on visits to private high schools. More specifically, we investigate how visits to private high schools by public research universities compare to visits to private high schools by selective private universities.

@@

Analyses of recruiting visits to private high schools can contribute novel insights about the "privatization" of public research universities. Within the scholarly literature on privatization, one strand of scholarship conceptualizes privatization as a process whereby declines in government funding cause public institutions to grow alternative revenue sources, particularly revenue from tuition and research commercialization [@RN4543; @RN2101]. Of relevance to this chapter, public research universities responded to declines in state appropriations by growing enrollment from nonresident students [@RN3753], who typically pay two- to three-times more than resident students. A second strand within the privatization literature conceptualizes privatization as a process whereby public institutions become more similar -- in terms of characteristics or behavior -- to for-profit or private non-profit organizations [@RN4543]. A weakness of this literature is that few empirical studies analyze public institutions on a behavior associated with private institutions.

@@

>>>>

---
title: "The Private School Network: Recruiting Visits to Private High Schools by Public and Private Universities"
author: 
  - Ozan Jaquette
  - Crystal Han
  - Irma Castañeda
bibliography: ./../other/burd_recruiting_chapter.bib
output: 
  bookdown::pdf_document2:
    toc: FALSE
    pandoc_args: !expr rmdfiltr::add_wordcount_filter(rmdfiltr::add_citeproc_filter(args = NULL))
csl: ./../other/apa.csl
urlcolor: blue
header-includes:
      - \usepackage{pdflscape}
      - \usepackage{geometry}
      - \usepackage{floatrow}
      - \usepackage{setspace}
      #- \doublespacing
      - \floatsetup{capposition=top}
fontsize: 12pt
---

@@

\setstretch{1.5}

@@

Analyses of recruiting visits to private high schools can contribute novel insights about the "privatization" of public research universities.

Within the scholarly literature on privatization, one strand of scholarship conceptualizes privatization as a process whereby declines in government funding cause public institutions to grow alternative revenue sources, particularly revenue from tuition and research commercialization [@RN4543; @RN2101]. Of relevance to this chapter, public research universities responded to declines in state appropriations by growing enrollment from nonresident students [@RN3753], who typically pay two- to three-times more than resident students. A second strand within the privatization literature conceptualizes privatization as a process whereby public institutions become more similar -- in terms of characteristics or behavior -- to for-profit or private non-profit organizations [@RN4543]. A weakness of this literature is that few empirical studies analyze public institutions on a behavior associated with private institutions. [FURTHER, FEW STUDIES COMPARE THE BEHAVIOR OF PUBLIC UNIVERSITIES TO THAT OF PRIVATE UNIVERSITIES ON THAT BEHAVIOR.]

Prior to the pandemic, admissions counselors from colleges and universities traveled to local high schools each year in order to meet prospective students and maintain relationships with guidance counselors at feeder schools [@RN3519]. @RN4758 analyzed off-campus recruiting visits made in 2017 by 15 public research universities. Surprisingly, 12 of the 15 universities made more visits to *out-of-state* high schools than *in-state* high schools. These out-of-state visits focused on public schools in affluent, predominantly white communities. Additionally, most public research universities in our sample made a disproportionate number of visits to out-of-state private high schools. For the University of California-Berkeley, for example, 155 (43%) of the 360 out-of-state high school visits were to private high schools. For the University of Colorado-Boulder, 321 (36%) of the 901 out-of-state high school visits were to private high schools.

@@

Whereas @RN4758 primarily analyzed visits to public high schools, this chapter focuses on visits to private high schools. More specifically, we investigate how visits to private high schools by public research universities compare to visits to private high schools by selective private universities.

@@

Little is known about recruiting at private high schools by public universities. Therefore, we conduct descriptive analyses that seek to identify basic patterns and relationships. Because a recruiting visit to the same private high school is an indicator of a relationship between two actors, we also utilize social network methods in order to compare the recruiting networks of public universities to those of private universities. Analyses are informed by the following three research questions:

<!--  ORIGINAL INTRODUCTION

Prior to the pandemic, admissions counselors from colleges and universities traveled to local high schools each year in order to meet prospective students and maintain relationships with guidance counselors at feeder schools [@RN3519]. @RN4758 analyzed off-campus recruiting visits made in 2017 by 15 public research universities. Surprisingly, 12 of the 15 universities made more visits to *out-of-state* high schools than *in-state* high schools. These out-of-state visits focused on public schools in affluent, predominantly white communities. Additionally, most public research universities in our sample made a disproportionate number of visits to out-of-state private high schools. For the University of California-Berkeley, for example, 155 (43%) of the 360 out-of-state high school visits were to private high schools. For the University of Colorado-Boulder, 321 (36%) of the 901 out-of-state high school visits were to private high schools.

Whereas @RN4758 primarily analyzed visits to public high schools, this chapter focuses on visits to private high schools. More specifically, we investigate how visits to private high schools by public research universities compare to visits to private high schools by selective private universities.

The mission of selective private colleges and universities is more squarely associated with social reproduction rather than social mobility. Historically, these institutions depend on patronage from philanthropic interests -- in connection with business and religious interests -- and on wealthy households who provide tuition revenue and donations [@RN4517]. In turn, selective private institutions contribute to social reproduction by providing access to their patrons. The "Chetty" data on parental income by college show that the overwhelming majority of students at selective private institutions have parental income in the top quintile [@RN4722; @RN4723]. A significant caveat is that religiously affiliated universities may prioritize enrollment opportunities for students of the same faith regardless of wealth, but even here the mission is to benefit families from a particular cultural status group.

Analyses of recruiting visits to private high schools can contribute novel insights about the "privatization" of public research universities. Within the scholarly literature on privatization, one strand of scholarship conceptualizes privatization as a process whereby declines in government funding cause public institutions to grow alternative revenue sources, particularly revenue from tuition and research commercialization [@RN4543; @RN2101]. Of relevance to this chapter, public research universities responded to declines in state appropriations by growing enrollment from nonresident students [@RN3753], who typically pay two- to three-times more than resident students. A second strand within the privatization literature conceptualizes privatization as a process whereby public institutions become more similar -- in terms of characteristics or behavior -- to for-profit or private non-profit organizations [@RN4543]. A weakness of this literature is that few empirical studies analyze public institutions on a behavior associated with private institutions. [FURTHER, FEW STUDIES COMPARE THE BEHAVIOR OF PUBLIC UNIVERSITIES TO THAT OF PRIVATE UNIVERSITIES ON THAT BEHAVIOR.]

Therefore, this chapter compares public research universities to selective private universities on the dimension of recruiting visits to private high schools, a behavior associated with the social reproduction orientation of selective private universities. Our primary goal is to learn about public research universities, using the behavior of private universities as a benchmark to develop more thoughtful insights about public research universities.  We analyze recruiting visits made in the 2017 calendar year by a sample of 15 public research universities and 14 selective private universities.

@@

\pagebreak
\singlespacing
